---
title: "Exam1"
author: "Maria"
date: "2023-03-07"
output: 
  word_document: default
  html_document:
    keep_md: yes
---

### Ex. 4

The ‘ggplot’ works in layers. A layer is a combination of data, stat and/or geom. Layering is the process of adding multiple layers (graphical elements) to a graph. You can add layers using the ‘+’ operator in ggplot and it is placed at the end of each line of the code containing a layer. For example, we can add points ‘geom_point()’ and lines ‘geom_smooth()’ to a plot by adding them (‘+’) as separate layers.
Scales control the mapping between data and aesthetics, for instance the size, type, or color of points in a plot. We can customize the scales using the ‘scale_ function’ in ggplot. For example, the ‘scale_color_gradient()’ function changes the color scale.
Themes customize the overall appearance of the plot, such as titles, background color, gridlines, legends, etc. You can customize the theme using the ‘theme_ function’ in ggplot. For example, ‘theme_grey()’ results in theme with a grey background and ‘theme_bw()’ a theme with a white background.
A facet breaks a plot into panels. Faceting is the process of dividing the data into subsets, based on the values of a categorical variable, then generates separated plots for each subset. You can create facets using the ‘facet_ wrap(~variablename)’ in ggplot.


```{r}

library (ggplot2)
plantpop <- read.csv("PlantPop.csv")

##Layering##

ggplot(plantpop, aes (x=Date, y=Yield, color=Cult)) +
  geom_point() +
  geom_smooth (method = lm, se = FALSE) +
  ylab ("Yield (Bu/Acre)") +
  xlab ("Planting Date")

##Scales##

ggplot(plantpop, aes (x=Date, y=Yield)) +
  geom_point(aes(color=Date)) +
  geom_smooth (method = lm, se = FALSE) +
  ylab ("Yield (Bu/Acre)") +
  xlab ("Planting Date")

ggplot (plantpop, aes (x=Date, y=Yield)) + 
  stat_summary(fun=mean, geom="bar", position = "dodge", aes(color=Date, fill=Date)) +
  stat_summary(fun.data = mean_se, geom = "errorbar", position = position_dodge(width=0.9), width = 0.5) +
  ylab ("Yield (Bu/Acre)") +
  xlab ("Planting Date")

##Themes##

ggplot (plantpop, aes (x=Date, y=Yield)) + 
  stat_summary(fun=mean, geom="bar", position = "dodge", aes(color=Date, fill=Date)) +
  stat_summary(fun.data = mean_se, geom = "errorbar", position = "dodge", width = 0.5) +
  ylab ("Yield (Bu/Acre)") +
  xlab ("Planting Date") +
  theme_classic()

ggplot (plantpop, aes (x=Date, y=Yield)) + 
  stat_summary(fun=mean, geom="bar", position = "dodge", aes(color=Date, fill=Date)) +
  stat_summary(fun.data = mean_se, geom = "errorbar", position = "dodge", width = 0.5) +
  ylab ("Yield (Bu/Acre)") +
  xlab ("Planting Date") +
  theme_gray()

##Facets##

ggplot (plantpop, aes (x=Date, y=Yield)) + 
  stat_summary(fun=mean, geom="bar", position = "dodge", aes(color=Date, fill=Date)) +
  stat_summary(fun.data = mean_se, geom = "errorbar", position = "dodge", width = 0.5) +
  ylab ("Yield (Bu/Acre)") +
  xlab ("Planting Date") +
  theme_classic() +
  facet_wrap(~Cult, scale = "free")

ggplot (plantpop, aes (x=Date, y=Yield)) + 
  stat_summary(fun=mean, geom="bar", position = "dodge", aes(color=Date, fill=Date)) +
  stat_summary(fun.data = mean_se, geom = "errorbar", position = "dodge", width = 0.5) +
  ylab ("Yield (Bu/Acre)") +
  xlab ("Planting Date") +
  theme_classic() +
  facet_wrap(~Cult*Density, scale = "free")


```

### Ex. 5

5.	Vector, matrix, and dataframe are all data structure that can be used in R. 
The vector is the most basic R data objects. It is a one-dimensional display of elements of the same data type (numeric, character, or logical). It can be thought of a list of elements saved together as an object. Vectors are used to create dataframes and matrixes.
A matrix, differently from the vector, is a two-dimensional array of elements. It is an atomic object that can contain only one type of data class (numeric, character, or logical). It is created using a vector input and it has fixed number of columns and rows. It can be thought of as a table of elements. A matrix can be treated as a vector if we have a single column or row.
A dataframe is very similar to a matrix, it is also a two-dimensional array of elements, but differently it can contain multiple data types together. A dataframe is a list of vectors with equal length where each columns can be a different data types. It can be thought of as a spreadsheet.

```{r}
subset1 <- ToothGrowth[ToothGrowth$supp == "VC", ]
subset1
subset1 <- subset(ToothGrowth, (supp == "VC"))
subset1
subset2 <- ToothGrowth[ToothGrowth$supp == "VC" & ToothGrowth$dose == 0.5, ]
subset2
subset2 <- subset(ToothGrowth, (supp == "VC" & dose == 0.5))
subset2
subset3 <- ToothGrowth[ToothGrowth$supp == "VC" & ToothGrowth$dose == 0.5, c("len")]
subset3
subset3 <- subset(ToothGrowth, supp == "VC" & dose == 0.5, select = "len")
subset3
```

